---
title: "demo-model-RDS"
author: "Greg Lenane"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
library(odbc)
library(DBI)
library(glue)
library(dplyr)
library(tidyr)
library(dbplyr)
library(lubridate)
library(ggplot2)
library(plotly)
library(scales)
library(openxlsx)

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, 
                      message = FALSE,
                      fig.width = 17, fig.height = 10)

# OAO_PRODUCTION DB connection
con_prod <- dbConnect(odbc(), "OAO Cloud DB Production")
# capacity modeling path
cap_dir <- "/SharedDrive/deans/Presidents/SixSigma/Project Based/System/Capacity Modeling/"

# excel wb function
add_to_wb <- function(df, sheetname, add_filter = FALSE) {
  # create sheet name
  sheet <- addWorksheet(wb, sheetname)
  
  setColWidths(wb, sheet, cols = 1:ncol(df), widths = "auto")
  
  # write data to sheet
  writeDataTable(wb,
                 x = df,
                 sheet = sheet,
                 keepNA = FALSE,
                 withFilter = add_filter)
  
  #create cell styles
  header_style <- createStyle(fgFill = "grey", halign = "left", 
                             textDecoration = "bold", fontColour = "#010101")
  body_style <- createStyle(border = "TopBottomLeftRight")
  
  # apply cell styles
  addStyle(wb, sheet, header_style, rows = 1, 
           cols = 1:ncol(df), gridExpand = TRUE, stack = TRUE)
  addStyle(wb, sheet, body_style, rows = 1:(nrow(df) + 1),
           cols = 1:ncol(df), gridExpand = TRUE, stack = TRUE)
  
}
```

```{r sinai colors, include=FALSE} 
# Sinai color scheme
mshs_colors <- c("#221F72", "#00AEFF", "#D80B8C", "#7F7F7F", "#000000", "#800080", "#FFFF00", "#CC0000", "#38761D", "#F39C12")
```

# Set Scenario Variables
```{r set scenario}

# set hospitals in scenario
hospitals <- list(
  "MSH",
  "MSM"
)

# set of services to be swapped in scenario
services <- list(
  "CARDIOLOGY",
  "CARDIOVASCULAR SURGERY"
)
```

# Data Processing
```{r dataset list}
datasets_processed <- list(
  "baseline" = readRDS(paste0(cap_dir, "Model Outputs/RDS/baseline_processed.RDS")),
  "scenario" = readRDS(paste0(cap_dir, "Model Outputs/RDS/scenario_processed.RDS")))
```

### IP Demand & Utilization
```{r ip}

# get billing descripions for IP categories
ip_mapping <- charge_cat %>% filter(CATEGORY %in% c("IP"))

# filter data down to IP bed charges in desired date range
ip_utilization <- lapply(datasets_processed, function(df) {
  df %>%
    filter(!is.na(EXTERNAL_NAME),
           BILLING_CAT_DESC %in% ip_mapping$BILLING_CAT_DESC) %>%
    group_by(FACILITY_MSX, SERVICE_GROUP, SERVICE_MONTH, MONTH_DAYS, SERVICE_DATE) %>%
    summarise(DAILY_DEMAND = sum(QUANTITY), .groups = "drop") %>%
    left_join(bed_cap, by = c("FACILITY_MSX" = "HOSPITAL", 
                              "SERVICE_GROUP" = "SERVICE_GROUP",
                              "SERVICE_MONTH" = "SERVICE_MONTH")) %>%
    mutate(UTILIZATION = DAILY_DEMAND/AVG_BED_CAPACITY,
           UTILIZATION_90 = case_when(
             UTILIZATION > .90 ~ TRUE,
             TRUE ~ FALSE),
           UTILIZATION_95 = case_when(
             UTILIZATION > .95 ~ TRUE,
             TRUE ~ FALSE))
})
```

### Lab & Radiology Demand
```{r lab & radiology}

# get billing descripions for lab and rad categories
lab_rad_mapping <- charge_cat %>% filter(CATEGORY %in% c("LAB", "RAD"))

# filter down datasets to only include lab and radiology charges
lab_rad_demand <- lapply(datasets_processed, function(df) {
  df %>%
    filter(BILLING_CAT_DESC %in% lab_rad_mapping$BILLING_CAT_DESC) %>%
    left_join(lab_rad_mapping, by = c("BILLING_CAT_DESC"="BILLING_CAT_DESC")) %>%
    group_by(FACILITY_MSX, SERVICE_GROUP, SERVICE_MONTH, SERVICE_DATE, CATEGORY) %>%
    summarise(COUNT = n())
})
```

# Scenario Outputs
## Inpatient
### IP Demand & Utilization Comparison
```{r ip demand and utilization comparison}

# compare utilization of baseline and scenario at daily level
ip_results_daily <- ip_utilization[["baseline"]] %>%
  full_join(ip_utilization[["scenario"]],
            by = c("FACILITY_MSX"="FACILITY_MSX",
                   "SERVICE_GROUP"="SERVICE_GROUP",
                   "SERVICE_MONTH"="SERVICE_MONTH",
                   "SERVICE_DATE"="SERVICE_DATE",
                   "AVG_BED_CAPACITY"="AVG_BED_CAPACITY"),
            suffix = c("_BASELINE", "_SCENARIO")) %>%
  filter(FACILITY_MSX %in% hospitals)

# aggregate comparison at monthly level
ip_results_monthly <- ip_results_daily %>%
  group_by(FACILITY_MSX, SERVICE_GROUP, SERVICE_MONTH, AVG_BED_CAPACITY) %>%
  summarise(AVG_DAILY_DEMAND_BASELINE = mean(DAILY_DEMAND_BASELINE),
            AVG_DAILY_DEMAND_SCENARIO = mean(DAILY_DEMAND_SCENARIO),
            AVG_UTILIZATION_BASELINE = mean(UTILIZATION_BASELINE),
            AVG_UTILIZATION_SCENARIO = mean(UTILIZATION_SCENARIO),
            PERCENT_90_BASELINE = mean(UTILIZATION_90_BASELINE),
            PERCENT_90_SCENARIO = mean(UTILIZATION_90_SCENARIO),
            PERCENT_95_BASELINE = mean(UTILIZATION_95_BASELINE),
            PERCENT_95_SCENARIO = mean(UTILIZATION_95_SCENARIO)) %>%
  mutate(across(where(is.numeric), \(x) coalesce(x, 0))) %>%
  mutate(AVG_UTILIZATION_SCENARIO = if_else(AVG_UTILIZATION_SCENARIO == 0, Inf, AVG_UTILIZATION_SCENARIO))

# aggregate comparison at total level
ip_results_total <- ip_results_monthly %>%
  group_by(FACILITY_MSX, SERVICE_GROUP) %>% 
  summarise(AVG_DAILY_DEMAND_BASELINE = mean(AVG_DAILY_DEMAND_BASELINE),
            AVG_DAILY_DEMAND_SCENARIO = mean(AVG_DAILY_DEMAND_SCENARIO),
            AVG_UTILIZATION_BASELINE =  mean(AVG_UTILIZATION_BASELINE),
            AVG_UTILIZATION_SCENARIO =  mean(AVG_UTILIZATION_SCENARIO),
            AVG_PERCENT90_BASELINE = mean(PERCENT_90_BASELINE),
            AVG_PERCENT90_SCENARIO = mean(PERCENT_90_SCENARIO),
            AVG_PERCENT95_BASELINE = mean(PERCENT_95_BASELINE),
            AVG_PERCENT95_SCENARIO = mean(PERCENT_95_SCENARIO),
            AVG_BED_CAPACITY = mean(AVG_BED_CAPACITY),
            AVG_VARIANCE_BASELINE = AVG_DAILY_DEMAND_BASELINE - AVG_BED_CAPACITY,
            AVG_VARIANCE_SCENARIO = AVG_DAILY_DEMAND_SCENARIO - AVG_BED_CAPACITY)
``` 

### IP Utilization Output
```{r utilization output}

# create table for output
ip_utilization_output <- ip_results_total %>%
  pivot_wider(id_cols = SERVICE_GROUP,
              names_from = FACILITY_MSX,
              values_from = c(AVG_BED_CAPACITY,
                              AVG_DAILY_DEMAND_BASELINE,
                              AVG_DAILY_DEMAND_SCENARIO,
                              AVG_UTILIZATION_BASELINE,
                              AVG_UTILIZATION_SCENARIO,
                              AVG_PERCENT90_BASELINE,
                              AVG_PERCENT90_SCENARIO,
                              AVG_PERCENT95_BASELINE,
                              AVG_PERCENT95_SCENARIO,
                              AVG_VARIANCE_BASELINE,
                              AVG_VARIANCE_SCENARIO)) %>%
  select(SERVICE_GROUP, contains(hospitals[[1]]), contains(hospitals[[2]]))
```

## Lab & Radiology
### Lab & Radiology Demand Comparison
```{r lab & radiology demand comparison}

# compare lab and radiology demand at daily level
lab_rad_results_daily <- lab_rad_demand[["baseline"]] %>%
  left_join(lab_rad_demand[["scenario"]],
            by = c("FACILITY_MSX"="FACILITY_MSX",
                   "SERVICE_GROUP"="SERVICE_GROUP",
                   "SERVICE_MONTH"="SERVICE_MONTH",
                   "SERVICE_DATE"="SERVICE_DATE",
                   "CATEGORY"="CATEGORY"),
            suffix = c("_BASELINE", "_SCENARIO")) %>%
  filter(FACILITY_MSX %in% hospitals)
```

### Lab & Rad Output
```{r lab & rad demand output}

lab_rad_output <- lab_rad_results_daily %>%
  group_by(FACILITY_MSX, SERVICE_MONTH, CATEGORY) %>%
  summarise(COUNT_BASELINE = sum(COUNT_BASELINE, na.rm = T),
            COUNT_SCENARIO = sum(COUNT_SCENARIO, na.rm = T)) %>%
  pivot_wider(id_cols = SERVICE_MONTH,
              names_from = c(CATEGORY, FACILITY_MSX),
              values_from = c(COUNT_BASELINE, 
                              COUNT_SCENARIO)) %>%
  mutate()
  select(SERVICE_MONTH, contains("LAB"), contains("RAD")) %>%
  select(SERVICE_MONTH, contains(hospitals[[1]]), contains(hospitals[[2]]))


```

### Excel WB
```{r excel workbook}

# create excel workbook for model outputs
wb <- createWorkbook()

# add sheet with IP Utilization Output
add_to_wb(df = ip_utilization_output,
          sheetname = "IP Utilization Comparison")
# add sheet with daily IP Stats
add_to_wb(df = ip_results_daily,
          sheetname = "IP Stats Daily",
          add_filter = TRUE)
# add sheet with daily Lab & Rad Stats
add_to_wb(df = lab_rad_results_daily,
          sheetname = "IP Stats Daily",
          add_filter = TRUE)

saveWorkbook(wb, 
             file = paste0(cap_dir, "Model Outputs/OUTPUT_",
                           hospitals[[1]], services[[1]], "-",
                           hospitals[[2]], services[[2]],"_", Sys.Date(), ".xlsx"),
             overwrite = TRUE )
```

## Visualizations
### Baseline vs. Scenario IP Utilization
```{r baseline vs. scenario}

# get list of service groups effected by scenario
service_change <- ip_results_total %>%
  filter(AVG_UTILIZATION_BASELINE != AVG_UTILIZATION_SCENARIO) %>%
  ungroup() %>%
  distinct(SERVICE_GROUP) %>%
  pull()

# prep data from graph
results_long <- ip_results_monthly %>%
  filter(SERVICE_GROUP %in% service_change) %>%
  pivot_longer(
    cols = c(AVG_UTILIZATION_BASELINE, AVG_UTILIZATION_SCENARIO),
    names_to = "Scenario",
    values_to = "Utilization") %>%
  mutate(
    Scenario = case_when(
      Scenario == "AVG_UTILIZATION_BASELINE" ~ "Baseline",
      Scenario == "AVG_UTILIZATION_SCENARIO" ~ "Scenario"),
    # dummy 95% utilization for services introduced by scenario
    Utilization = case_when(
      Scenario == "Scenario" & is.na(Utilization) ~ 1,
      TRUE ~ Utilization))

# Plot
ggplot(results_long, aes(x = SERVICE_MONTH, y = Utilization, fill = Scenario)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_hline(yintercept = .95, color = "black", linetype = "solid", linewidth = 0.7) +
  scale_y_continuous(limits = c(0, 1.5), labels = scales::percent) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") +
  scale_fill_manual(values = mshs_colors) + 
  facet_grid(FACILITY_MSX ~ SERVICE_GROUP) +
  labs(title = "Utilization Comparison",
       x = "Service Month",
       y = "Utilization",
       fill = "Scenario") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
  panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5),
    strip.text = element_text(face = "bold"))
```

### Daily Scenario IP Utilization
```{r scenario utilization}

# Plot
ggplot(ip_results_daily, aes(
  x = SERVICE_DATE,
  y = UTILIZATION_SCENARIO,
  color = SERVICE_GROUP,
  group = SERVICE_GROUP,
  text = paste0("Service Group: ", SERVICE_GROUP,
                "<br>Utilization: ", round(UTILIZATION_SCENARIO * 100, 1), "%"))
) +
geom_smooth(se = FALSE, method = "loess", span = 0.2, linewidth = 1.2) +
geom_hline(yintercept = 0.95, color = "black", linewidth = 1.2) +
facet_wrap(~ FACILITY_MSX) +
scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 1.3)) +
scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") +
scale_color_manual(name = "Service Group", values = mshs_colors) +   
labs(
  title = "Utilization Over Time",
  x = "Service Date",
  y = "Utilization (Scenario)"
) +
theme_minimal() +
theme(
  legend.position = "bottom",
  axis.text.x = element_text(angle = 45, hjust = 1),
  strip.text = element_text(face = "bold"),
  panel.border = element_rect(color = "black", fill = NA)
)
```

### Daily Lab & Radiology Demand
```{r lab & rad daily trend}

# visualization
results_long <- lab_rad_results_daily %>%
  filter(FACILITY_MSX %in% hospitals) %>%
  group_by(FACILITY_MSX, SERVICE_DATE, CATEGORY) %>%
  summarise(COUNT_BASELINE = sum(COUNT_BASELINE),
            COUNT_SCENARIO = sum(COUNT_SCENARIO)) %>%
  pivot_longer(
    cols = c(COUNT_BASELINE, COUNT_SCENARIO),
    names_to = "type",
    values_to = "count"
  ) %>%
  group_by(FACILITY_MSX, SERVICE_DATE, CATEGORY)

# make the plot
ggplot(results_long, aes(x = SERVICE_DATE, y = count, color = type, group = type)) +
  geom_line(size = 1) +
  geom_point() +
  facet_grid(CATEGORY ~ FACILITY_MSX, scales = "free_y") +
  scale_color_manual(values = mshs_colors) +
  labs(
    title = "Baseline vs. Scenario Lab & Radiology Counts",
    x = "Date",
    y = "Count",
    color = "Type"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

# Quality Checks
### Epic Encounter Check
```{r encounter check}

# OAO_PRODUCTION DB connection
con_prod <- dbConnect(odbc(), "OAO Cloud DB Production")

msx_encounters <- glue(
  "SELECT DISTINCT EPIC_CSN
   FROM MSX_IP_OUTPUT
   WHERE ADMIT_DT_SRC >= DATE '2024-01-01' AND 
         ADMIT_DT_SRC <= DATE '2024-12-31' AND 
         FACILITY_MSX <> 'MSSN';")

oe_encounters <- glue(
  "WITH ip AS (
       SELECT *
       FROM MSX_IP_OUTPUT
       WHERE ADMIT_DT_SRC >= DATE '2024-01-01'
         AND ADMIT_DT_SRC <= DATE '2024-12-31'
         AND FACILITY_MSX <> 'MSSN'
   ),
   charge AS (
       select *
       from OE_CHARGE_DETAIL
       where PRIM_ENC_CSN_ID IN (SELECT EPIC_CSN FROM ip))
  select distinct PRIM_ENC_CSN_ID from charge"
)

# create, execute and clear result for MSX IP Encounters
msx_send <- dbSendQuery(con_prod, msx_encounters)
msx <- dbFetch(msx_send)
dbClearResult(msx_send)

# create, execute and clear result for MSX IP Encounters
oe_send <- dbSendQuery(con_prod, oe_encounters)
oe <- dbFetch(oe_send)
dbClearResult(oe_send)

percent_of_encounters <- 
  (length(msx$EPIC_CSN) - length(setdiff(msx$EPIC_CSN, oe$PRIM_ENC_CSN_ID)))/length(msx$EPIC_CSN) 

print(percent_of_encounters)

```

### Bed Charges Over Time
```{r bed charges trend}
query <- glue(
  "WITH ip AS (
       SELECT *
       FROM MSX_IP_OUTPUT
       WHERE ADMIT_DT_SRC >= DATE '2024-06-01' AND 
             ADMIT_DT_SRC <= DATE '2024-12-31' AND
             FACILITY_MSX <> 'MSSN'
   ),
   charge AS (
       select *
       from OE_CHARGE_DETAIL
       where PRIM_ENC_CSN_ID IN (SELECT EPIC_CSN FROM ip) AND
             BILLING_CAT_DESC = 'BED CHARGES'
   )
   select SERVICE_DATE from charge;"
)

# create, execute and clear result for MSX IP Charge Desctiptions
send <- dbSendQuery(con_prod, query)
datacheck <- dbFetch(send)
dbClearResult(send)

test <- datacheck %>%
  mutate(SERVICE_DATE = as.Date(paste0(
    substr(SERVICE_DATE, 1, 4), "-",
    substr(SERVICE_DATE, 5, 6), "-",
    substr(SERVICE_DATE, 7, 8)))) %>%
  #filter(SERVICE_DATE >= as.Date("2024-06-01")) %>%
  group_by(SERVICE_DATE) %>%
  summarise(count = n())

ggplot(test, aes(x = SERVICE_DATE, y = count)) +
  geom_line(color = "steelblue", linewidth = 1) +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b %Y"  # e.g., Jan 2025
  ) +
  labs(
    title = "Bed Charge Count Over Time",
    x = "Service Date",
    y = "Count"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```