---
title: "demo-model"
author: "Greg Lenane"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
library(odbc)
library(DBI)
library(glue)
library(dplyr)
library(tidyr)
library(dbplyr)
library(lubridate)
library(ggplot2)
library(plotly)
library(scales)
library(openxlsx)

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, 
                      message = FALSE,
                      fig.width = 17, fig.height = 10)

# OAO_PRODUCTION DB connection
con_prod <- dbConnect(odbc(), "OAO Cloud DB Production")
# capacity modeling path
cap_dir <- "/SharedDrive/deans/Presidents/SixSigma/Project Based/System/Capacity Modeling/"

# excel wb function
add_to_wb <- function(df, sheetname, add_filter = FALSE) {
  # create sheet name
  sheet <- addWorksheet(wb, sheetname)
  
  setColWidths(wb, sheet, cols = 1:ncol(df), widths = "auto")
  
  # write data to sheet
  writeDataTable(wb,
                 x = df,
                 sheet = sheet,
                 keepNA = FALSE,
                 withFilter = add_filter)
  
  #create cell styles
  header_style <- createStyle(fgFill = "grey", halign = "left", 
                             textDecoration = "bold", fontColour = "#010101")
  body_style <- createStyle(border = "TopBottomLeftRight")
  
  # apply cell styles
  addStyle(wb, sheet, header_style, rows = 1, 
           cols = 1:ncol(df), gridExpand = TRUE, stack = TRUE)
  addStyle(wb, sheet, body_style, rows = 1:(nrow(df) + 1),
           cols = 1:ncol(df), gridExpand = TRUE, stack = TRUE)
  
}
```

```{r sinai colors, include=FALSE} 
# Sinai color scheme
mshs_colors <- c("#221F72", "#00AEFF", "#D80B8C", "#7F7F7F", "#000000", "#800080", "#FFFF00", "#CC0000", "#38761D", "#F39C12")
```

# Set Scenario Variables
```{r set scenario}

# set hospitals in scenario
hospitals <- list(
  "MSH",
  "MSM"
)

# set of services to be swapped in scenario
services <- list(
  "CARDIOLOGY",
  "CARDIOVASCULAR SURGERY"
)
```

# Baseline Data
```{r read baseline data}

# read in bed capacity (tableau)
# List all CSV files in the directory
bed_cap_csv <- list.files(paste0(cap_dir, "Tableau Data/Bed Capacity/"),
                          pattern = "\\.csv$", full.names = TRUE)
bed_cap <- bed_cap_csv %>%
  map_dfr(~ read_csv(.x) %>% mutate(source_file = basename(.x))) %>%
  rename(HOSPITAL = Location,
         SERVICE_GROUP = `Service Group`,
         MEASURE = `Measure Names`,
         SERVICE_MONTH = `Day of Census Start`) %>%
  filter(MEASURE == 'Avg Total Beds') %>%
  mutate(
    HOSPITAL = case_when(
      HOSPITAL == "MOUNT SINAI BETH ISRAEL" ~ "MSBI",
      HOSPITAL == "MOUNT SINAI BROOKLYN" ~ "MSB",
      HOSPITAL == "MOUNT SINAI MORNINGSIDE" ~ "MSM",
      HOSPITAL == "MOUNT SINAI QUEENS" ~ "MSQ",
      HOSPITAL == "MOUNT SINAI WEST" ~ "MSW",
      HOSPITAL == "THE MOUNT SINAI HOSPITAL" ~ "MSH"),
    SERVICE_MONTH = mdy(SERVICE_MONTH)) %>%
  group_by(HOSPITAL, SERVICE_GROUP, SERVICE_MONTH) %>%
  summarise(AVG_BED_CAPACITY = sum(`Measure Values`, na.rm = TRUE))

# read in charge description categorization file
charge_cat <- readxl::read_xlsx(paste0(cap_dir, "MSX/BILLING_CAT_DESC.xlsx")) %>%
  filter(!is.na(CATEGORY))
charge_desc <- paste(charge_cat$BILLING_CAT_DESC, 
                              collapse = "\',\'")

# pull in all encounters and their charges. join with service grouping for IP unit type
baseline_query <- glue(
  "WITH ip AS (
       SELECT *
       FROM MSX_IP_OUTPUT
       WHERE ADMIT_DT_SRC >= DATE '2024-05-01'
         AND ADMIT_DT_SRC <= DATE '2025-01-31'
         AND FACILITY_MSX <> 'MSSN'
   ),
   charge AS (
       select *
       from OE_CHARGE_DETAIL
       where PRIM_ENC_CSN_ID IN (SELECT EPIC_CSN FROM ip) AND
             BILLING_CAT_DESC in (\'{charge_desc}\')
   ),
   service_group AS (
       select *
       from DASHBD_USER.CLARITY_DEP_REF
   ),
   final as (
   SELECT ip.ENCOUNTER_NO,
          ip.FACILITY_MSX,
          ip.DSCH_DT_SRC,
          ip.DSCH_TIME_SRC,
          ip.ADMIT_DT_SRC,
          ip.ADMIT_TIME_SRC,
          ip.LOS_NO_SRC,
          ip.PRINCIPAL_SURGEON_CD_SRC,
          ip.PRINCIPAL_SURGEON_NAME_MSX,
          ip.MSDRG_CD_SRC,
          ip.MSDRG_DESC_MSX,
          ip.ADMIT_SOURCE_CD_MSX,
          ip.ADMIT_SOURCE_DESC_MSX,
          ip.P_AVG_LOS_MSDRG,
          ip.VERITY_DEPT_CD_SRC,
          ip.VERITY_DEPT_DESC_SRC,
          ip.VERITY_DIV_CD_SRC,
          ip.VERITY_DIV_DESC_SRC,
          charge.FACILITY_ABBR,
          charge.COST_CENTER_C,
          charge.CHARGE_C,
          charge.EPIC_DEPT_ID,
          charge.EPIC_DEPT_NAME,
          charge.QUANTITY,
          charge.SERVICE_DATE,
          charge.CPT_HCPCS_C,
          charge.BILLING_CAT_C,
          charge.BILLING_CAT_DESC,
          charge.RPT_GRP_NINETEEN_C,
          charge.RPT_GRP_NINETEEN_DESC,
          charge.NEW_COST_CENTER_C,
          charge.NEW_GL_COMPONENT,
          service_group.EXTERNAL_NAME,
          service_group.SERVICE_GROUP,
          service_group.RPT_GRP_TWENTYTHREE
   FROM ip
   LEFT JOIN charge
     ON ip.EPIC_CSN = charge.PRIM_ENC_CSN_ID
   LEFT JOIN service_group
     ON charge.EPIC_DEPT_ID = service_group.DEPARTMENT_ID
   ORDER BY ip.EPIC_CSN
   )
   select * from final;")

# create, execute and clear result for MSX IP Charge Desctiptions
baseline_send <- dbSendQuery(con_prod, baseline_query)
baseline <- dbFetch(baseline_send)
dbClearResult(baseline_send)

baseline <- baseline %>%
  mutate(
    FACILITY_MSX = case_when(
      FACILITY_MSX == 'STL' ~ 'MSM',
      FACILITY_MSX == 'RVT' ~ 'MSW',
      FACILITY_MSX == 'BIB' ~ 'MSB',
      FACILITY_MSX == 'BIP' ~ 'MSBI',
      TRUE ~ FACILITY_MSX))
```

# Scenario Data
```{r scenario data}

# create new dummy hospital column to represent new location for scenario population
scenario <- baseline %>%
  mutate(
    FACILITY_MSX = case_when(
      FACILITY_MSX == hospitals[[1]] & VERITY_DEPT_DESC_SRC == services[[2]] ~ "MSM",
      FACILITY_MSX == hospitals[[2]] & VERITY_DEPT_DESC_SRC == services[[1]] ~ "MSH",
      TRUE ~ FACILITY_MSX))
```

# Data Processing
```{r dataset list}
# place datasets in list to iterate through
datasets <- list(
  "baseline" = baseline,
  "scenario" = scenario
)

# preprocess raw datasets
datasets_processed <- lapply(datasets, function(df) {
  df %>%
    mutate(
      SERVICE_DATE = as.Date(paste0(
        substr(SERVICE_DATE, 1, 4), "-",
        substr(SERVICE_DATE, 5, 6), "-",
        substr(SERVICE_DATE, 7, 8))),
      SERVICE_MONTH = floor_date(SERVICE_DATE, unit = "month"),
      MONTH_DAYS = days_in_month(SERVICE_MONTH)) %>%
    filter(SERVICE_DATE <= as.Date("2024-12-31"),
           SERVICE_DATE >= as.Date("2024-06-01"))
})

# remove original dataframes to free up space
rm(baseline, scenario, datasets)
```

### IP Demand & Utilization
```{r ip}

# get billing descripions for IP categories
ip_mapping <- charge_cat %>% filter(CATEGORY %in% c("IP"))

# filter data down to IP bed charges in desired date range
ip_utilization <- lapply(datasets_processed, function(df) {
  df %>%
    filter(!is.na(EXTERNAL_NAME),
           BILLING_CAT_DESC %in% ip_mapping$BILLING_CAT_DESC) %>%
    group_by(FACILITY_MSX, SERVICE_GROUP, SERVICE_MONTH, MONTH_DAYS, SERVICE_DATE) %>%
    summarise(DAILY_DEMAND = sum(QUANTITY), .groups = "drop") %>%
    left_join(bed_cap, by = c("FACILITY_MSX" = "HOSPITAL", 
                              "SERVICE_GROUP" = "SERVICE_GROUP",
                              "SERVICE_MONTH" = "SERVICE_MONTH")) %>%
    mutate(UTILIZATION = DAILY_DEMAND/AVG_BED_CAPACITY,
           UTILIZATION_90 = case_when(
             UTILIZATION > .90 ~ TRUE,
             TRUE ~ FALSE),
           UTILIZATION_95 = case_when(
             UTILIZATION > .95 ~ TRUE,
             TRUE ~ FALSE))
})
```

### Lab & Radiology Demand
```{r lab & radiology}

# get billing descripions for lab and rad categories
lab_rad_mapping <- charge_cat %>% filter(CATEGORY %in% c("LAB", "RAD"))

# filter down datasets to only include lab and radiology charges
lab_rad_demand <- lapply(datasets_processed, function(df) {
  df %>%
    filter(BILLING_CAT_DESC %in% lab_rad_mapping$BILLING_CAT_DESC) %>%
    left_join(lab_rad_mapping, by = c("BILLING_CAT_DESC"="BILLING_CAT_DESC")) #%>%
    #group_by(FACILITY_MSX, SERVICE_GROUP, SERVICE_MONTH, SERVICE_DATE, CATEGORY) %>%
    #summarise(COUNT = n())
})
```

# Scenario Outputs
### IP Demand & Utilization Comparison
```{r}

# compare utilization of baseline and scenario at daily level
results_comparison_daily <- ip_utilization[["baseline"]] %>%
  full_join(ip_utilization[["scenario"]],
            by = c("FACILITY_MSX"="FACILITY_MSX",
                   "SERVICE_GROUP"="SERVICE_GROUP",
                   "SERVICE_MONTH"="SERVICE_MONTH",
                   "SERVICE_DATE"="SERVICE_DATE",
                   "AVG_BED_CAPACITY"="AVG_BED_CAPACITY"),
            suffix = c("_BASELINE", "_SCENARIO")) %>%
  filter(FACILITY_MSX %in% hospitals)

# aggregate comparison at monthly level
results_comparison_monthly <- results_comparison_daily %>%
  group_by(FACILITY_MSX, SERVICE_GROUP, SERVICE_MONTH, AVG_BED_CAPACITY) %>%
  summarise(AVG_DAILY_DEMAND_BASELINE = mean(DAILY_DEMAND_BASELINE),
            AVG_DAILY_DEMAND_SCENARIO = mean(DAILY_DEMAND_SCENARIO),
            AVG_UTILIZATION_BASELINE = mean(UTILIZATION_BASELINE),
            AVG_UTILIZATION_SCENARIO = mean(UTILIZATION_SCENARIO),
            PERCENT_90_BASELINE = mean(UTILIZATION_90_BASELINE),
            PERCENT_90_SCENARIO = mean(UTILIZATION_90_SCENARIO),
            PERCENT_95_BASELINE = mean(UTILIZATION_95_BASELINE),
            PERCENT_95_SCENARIO = mean(UTILIZATION_95_SCENARIO)) %>%
  mutate(across(where(is.numeric), \(x) coalesce(x, 0))) %>%
  mutate(AVG_UTILIZATION_SCENARIO = if_else(AVG_UTILIZATION_SCENARIO == 0, Inf, AVG_UTILIZATION_SCENARIO))

# aggregate comparison at total level
results_comparison_total <- results_comparison_monthly %>%
  group_by(FACILITY_MSX, SERVICE_GROUP) %>% 
  summarise(AVG_DAILY_DEMAND_BASELINE = mean(AVG_DAILY_DEMAND_BASELINE),
            AVG_DAILY_DEMAND_SCENARIO = mean(AVG_DAILY_DEMAND_SCENARIO),
            AVG_UTILIZATION_BASELINE =  mean(AVG_UTILIZATION_BASELINE),
            AVG_UTILIZATION_SCENARIO =  mean(AVG_UTILIZATION_SCENARIO),
            AVG_PERCENT90_BASELINE = mean(PERCENT_90_BASELINE),
            AVG_PERCENT90_SCENARIO = mean(PERCENT_90_SCENARIO),
            AVG_PERCENT95_BASELINE = mean(PERCENT_95_BASELINE),
            AVG_PERCENT95_SCENARIO = mean(PERCENT_95_SCENARIO),
            AVG_BED_CAPACITY = mean(AVG_BED_CAPACITY),
            AVG_VARIANCE_BASELINE = AVG_DAILY_DEMAND_BASELINE - AVG_BED_CAPACITY,
            AVG_VARIANCE_SCENARIO = AVG_DAILY_DEMAND_SCENARIO - AVG_BED_CAPACITY)
``` 

### IP Utilization Output
```{r utilization output}

# create table for output
utilization_output <- results_comparison_total %>%
  pivot_wider(id_cols = SERVICE_GROUP,
              names_from = FACILITY_MSX,
              values_from = c(AVG_BED_CAPACITY,
                              AVG_DAILY_DEMAND_BASELINE,
                              AVG_DAILY_DEMAND_SCENARIO,
                              AVG_UTILIZATION_BASELINE,
                              AVG_UTILIZATION_SCENARIO,
                              AVG_PERCENT90_BASELINE,
                              AVG_PERCENT90_SCENARIO,
                              AVG_PERCENT95_BASELINE,
                              AVG_PERCENT95_SCENARIO,
                              AVG_VARIANCE_BASELINE,
                              AVG_VARIANCE_SCENARIO)) %>%
  select(SERVICE_GROUP, contains(hospitals[[1]]), contains(hospitals[[2]]))
```

### Excel WB
```{r excel workbook}

# create excel workbook for model outputs
wb <- createWorkbook()

# add sheet with IP Utilization Output
add_to_wb(df = utilization_output,
          sheetname = "IP Utilization Comparison")
# add sheet with daily IP Stats
add_to_wb(df = results_comparison_daily,
          sheetname = "IP Stats Daily",
          add_filter = TRUE)

saveWorkbook(wb, 
             file = paste0(cap_dir, "Model Outputs/OUTPUT_",
                           hospitals[[1]], services[[1]], "-",
                           hospitals[[2]], services[[2]],"_", Sys.Date(), ".xlsx"),
             overwrite = TRUE )
```

### Visualizations
#### Baseline vs. Scenario IP Utilization
```{r baseline vs. scenario}

# get list of service groups effected by scenario
service_change <- results_comparison_total %>%
  filter(AVG_UTILIZATION_BASELINE != AVG_UTILIZATION_SCENARIO) %>%
  ungroup() %>%
  distinct(SERVICE_GROUP) %>%
  pull()

# prep data from graph
results_long <- results_comparison_monthly %>%
  filter(SERVICE_GROUP %in% service_change) %>%
  pivot_longer(
    cols = c(AVG_UTILIZATION_BASELINE, AVG_UTILIZATION_SCENARIO),
    names_to = "Scenario",
    values_to = "Utilization") %>%
  mutate(
    Scenario = case_when(
      Scenario == "AVG_UTILIZATION_BASELINE" ~ "Baseline",
      Scenario == "AVG_UTILIZATION_SCENARIO" ~ "Scenario"),
    # dummy 95% utilization for services introduced by scenario
    Utilization = case_when(
      Scenario == "Scenario" & is.na(Utilization) ~ 1,
      TRUE ~ Utilization))

# Plot
ggplot(results_long, aes(x = SERVICE_MONTH, y = Utilization, fill = Scenario)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_hline(yintercept = .95, color = "black", linetype = "solid", linewidth = 0.7) +
  scale_y_continuous(limits = c(0, 1.5), labels = scales::percent) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") +
  scale_fill_manual(values = mshs_colors) + 
  facet_grid(FACILITY_MSX ~ SERVICE_GROUP) +
  labs(title = "Utilization Comparison",
       x = "Service Month",
       y = "Utilization",
       fill = "Scenario") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
  panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5),
    strip.text = element_text(face = "bold"))
```

#### Daily Scenario IP Utilization
```{r scenario utilization}

# Plot
ggplot(results_comparison_daily, aes(
  x = SERVICE_DATE,
  y = UTILIZATION_SCENARIO,
  color = SERVICE_GROUP,
  group = SERVICE_GROUP,
  text = paste0("Service Group: ", SERVICE_GROUP,
                "<br>Utilization: ", round(UTILIZATION_SCENARIO * 100, 1), "%"))
) +
geom_smooth(se = FALSE, method = "loess", span = 0.2, linewidth = 1.2) +
geom_hline(yintercept = 0.95, color = "black", linewidth = 1.2) +
facet_wrap(~ FACILITY_MSX) +
scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 1.3)) +
scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") +
scale_color_manual(name = "Service Group", values = mshs_colors) +   
labs(
  title = "Utilization Over Time",
  x = "Service Date",
  y = "Utilization (Scenario)"
) +
theme_minimal() +
theme(
  legend.position = "bottom",
  axis.text.x = element_text(angle = 45, hjust = 1),
  strip.text = element_text(face = "bold"),
  panel.border = element_rect(color = "black", fill = NA)
)
```

#### Unit Type Demand for Scenario
```{r unit type demand}

# # this plot answers question of, in the scenario situation what would be the unit type demand for the scenario group
# # filter down data needed for plot
# unit_type_demand <- ip_utilization[["scenario"]] %>%
#   filter(FACILITY_MSX %in% hospitals,
#          VERITY_DEPT_DESC_SRC %in% services)
# 
# # define facet labels
# facility_labels <- setNames(paste(hospitals, "-", services),hospitals)
# 
# ggplot(unit_type_demand, aes(x = SERVICE_GROUP,
#                              fill = FACILITY_MSX)) +
#   geom_bar() +  
#   geom_text(
#     stat = "count",
#     aes(label = after_stat(count)),
#     vjust = -0.5,
#     size = 3) +
#   facet_wrap(~ FACILITY_MSX,
#              labeller = labeller(FACILITY_MSX = facility_labels)) + 
#   scale_y_continuous(
#     expand = expansion(mult = c(0, 0.1))) +
#   scale_fill_manual(values = mshs_colors) +
#   labs(
#    x = "Service Group",
#    y = "Count",
#    title = "Count of Unit Type by Scenario",
#    fill = "Hospital") +
#   theme(
#    axis.text.x = element_text(angle = 45, hjust = 1),
#    panel.border = element_rect(color = "black", fill = NA, linewidth = 1))
```

# Quality Checks
### Epic Encounter Check
```{r encounter check}

# OAO_PRODUCTION DB connection
con_prod <- dbConnect(odbc(), "OAO Cloud DB Production")

msx_encounters <- glue(
  "SELECT DISTINCT EPIC_CSN
   FROM MSX_IP_OUTPUT
   WHERE ADMIT_DT_SRC >= DATE '2024-01-01' AND 
         ADMIT_DT_SRC <= DATE '2024-12-31' AND 
         FACILITY_MSX <> 'MSSN';")

oe_encounters <- glue(
  "WITH ip AS (
       SELECT *
       FROM MSX_IP_OUTPUT
       WHERE ADMIT_DT_SRC >= DATE '2024-01-01'
         AND ADMIT_DT_SRC <= DATE '2024-12-31'
         AND FACILITY_MSX <> 'MSSN'
   ),
   charge AS (
       select *
       from OE_CHARGE_DETAIL
       where PRIM_ENC_CSN_ID IN (SELECT EPIC_CSN FROM ip))
  select distinct PRIM_ENC_CSN_ID from charge"
)

# create, execute and clear result for MSX IP Encounters
msx_send <- dbSendQuery(con_prod, msx_encounters)
msx <- dbFetch(msx_send)
dbClearResult(msx_send)

# create, execute and clear result for MSX IP Encounters
oe_send <- dbSendQuery(con_prod, oe_encounters)
oe <- dbFetch(oe_send)
dbClearResult(oe_send)

percent_of_encounters <- 
  (length(msx$EPIC_CSN) - length(setdiff(msx$EPIC_CSN, oe$PRIM_ENC_CSN_ID)))/length(msx$EPIC_CSN) 

print(percent_of_encounters)

```

### Bed Charges Over Time
```{r bed charges trend}
query <- glue(
  "WITH ip AS (
       SELECT *
       FROM MSX_IP_OUTPUT
       WHERE ADMIT_DT_SRC >= DATE '2024-01-01' AND 
             ADMIT_DT_SRC <= DATE '2024-12-31' AND
             FACILITY_MSX <> 'MSSN'
   ),
   charge AS (
       select *
       from OE_CHARGE_DETAIL
       where PRIM_ENC_CSN_ID IN (SELECT EPIC_CSN FROM ip) AND
             BILLING_CAT_DESC = 'BED CHARGES'
   )
   select SERVICE_DATE from charge;"
)

# create, execute and clear result for MSX IP Charge Desctiptions
send <- dbSendQuery(con_prod, query)
datacheck <- dbFetch(send)
dbClearResult(send)

test <- datacheck %>%
  mutate(SERVICE_DATE = as.Date(paste0(
    substr(SERVICE_DATE, 1, 4), "-",
    substr(SERVICE_DATE, 5, 6), "-",
    substr(SERVICE_DATE, 7, 8)))) %>%
  filter(SERVICE_DATE >= as.Date("2024-06-01")) %>%
  group_by(SERVICE_DATE) %>%
  summarise(count = n())

ggplot(test, aes(x = SERVICE_DATE, y = count)) +
  geom_line(color = "steelblue", linewidth = 1) +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b %Y"  # e.g., Jan 2025
  ) +
  labs(
    title = "Bed Charge Count Over Time",
    x = "Service Date",
    y = "Count"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```





