---
title: "Unit Switch x LOS Index"
author: "Greg Lenane"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	fig.width = 9, 
	fig.height = 5)

library(odbc)
library(DBI)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(knitr)
library(kableExtra)
library(patchwork)
library(scales)
library(gridExtra)
library(readr)
library(hexbin)

# OAO_PRODUCTION DB connection
con_prod <- dbConnect(odbc(), "OAO Cloud DB Production")
```

# Data Setup
### Load Baseline
```{r load baseline, message=FALSE}
# baseline IPCAP Bedcharge data
baseline <- tbl(con_prod, "IPCAP_BEDCHARGES") %>% collect() %>%
  mutate(SERVICE_DATE = as.Date(SERVICE_DATE, format = "%Y%m%d")) %>%
  filter(ADMIT_DT_SRC >= as.Date("2025-01-01"),
         DSCH_DT_SRC <= as.Date("2025-12-31"),
         LOS_NO_SRC <= 100,
         !is.na(LOS_NO_SRC),
         !is.na(VIZ_EX_LOS),
         !(SERVICE_DESC_MSX %in% c("PSYCHIATRY", "REHABILITATION", "CHEMICAL DEPENDECY")))
```

### Encounters by Day
```{r encounter by day, message=FALSE}

# LOS for each care team (care team must have a service line) on IP units
encounter_by_day <- baseline %>%
  group_by(ENCOUNTER_NO, EPIC_CSN, FACILITY_MSX, LOC_NAME, LOS_NO_SRC, VIZ_EX_LOS, ADMIT_DT_SRC, 
           DSCH_DT_SRC, ADMIT_UNIT_DESC_MSX, DSCH_UNIT_DESC_MSX, ADMIT_TYPE_DESC_SRC,
           ATTENDING_VERITY_DIV_DESC, MSDRG_CD_SRC, MSDRG_DESC_MSX, 
           EPIC_DEPT_ID, EPIC_DEPT_NAME, EXTERNAL_NAME, SERVICE_GROUP, 
           SERVICE_DATE) %>%
  summarise(BED_CHARGES = sum(QUANTITY)) %>%
  mutate(BED_CHARGES = case_when(
    BED_CHARGES > 1 ~ 1,
    TRUE ~ BED_CHARGES)) %>%
  filter(BED_CHARGES != 0) %>%
  arrange(ENCOUNTER_NO, SERVICE_DATE) %>%
  mutate(
    LOS_INDEX = LOS_NO_SRC / VIZ_EX_LOS,
    EXTERNAL_NAME = case_when(
      is.na(EXTERNAL_NAME) ~ paste0("(NON IP) ", EPIC_DEPT_NAME),
      TRUE ~ EXTERNAL_NAME),
    SERVICE_GROUP = case_when(
      is.na(SERVICE_GROUP) ~ EXTERNAL_NAME,
      TRUE ~ SERVICE_GROUP))
```

### Unit Stays and Patient Identifiers
```{r encounter units, message=FALSE}

# break down each encounters stay into unit stays by number of days
encounter_unit_stays <- encounter_by_day %>%  
  arrange(ENCOUNTER_NO, SERVICE_DATE) %>%
  group_by(ENCOUNTER_NO) %>%
  mutate(
    UNIT_CHANGE = EXTERNAL_NAME != lag(EXTERNAL_NAME),
    DATE_GAP    = SERVICE_DATE != lag(SERVICE_DATE) + days(1),
    NEW_STAY    = is.na(lag(SERVICE_DATE)) | UNIT_CHANGE | DATE_GAP,
    STAY_ID     = cumsum(NEW_STAY),
         
    SERVICE_CHANGE = SERVICE_GROUP != lag(SERVICE_GROUP),
    NEW_SERVICE    = is.na(lag(SERVICE_DATE)) | SERVICE_CHANGE | DATE_GAP,
    SERVICE_ID     = cumsum(NEW_SERVICE))
```

### Unit Switch
```{r unit stays, message=FALSE}
# for each encounter get the total unit switches
encounter_unit_switch <- encounter_unit_stays %>%
  group_by(ENCOUNTER_NO, EPIC_CSN, FACILITY_MSX, ATTENDING_VERITY_DIV_DESC, MSDRG_CD_SRC, 
           MSDRG_DESC_MSX, ADMIT_TYPE_DESC_SRC, LOS_NO_SRC, VIZ_EX_LOS, LOS_INDEX) %>%
  summarise(UNIT_COUNT = max(STAY_ID),
            SERVICE_COUNT = max(SERVICE_ID)) %>%
  mutate(FLAG = UNIT_COUNT > SERVICE_COUNT)

head(encounter_unit_switch) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                font_size = 10,
                full_width = F)
```

# Visualizations
## Distributions
### Full Population - LOS Distribution
```{r population visualization}
ggplot(encounter_unit_switch, aes(x = UNIT_COUNT, y = LOS_NO_SRC)) +
  geom_jitter(aes(colour = factor(UNIT_COUNT)),
              width = .2) +
  labs(
    title = "Distribution of Unit Count and LOS Index",
    x = "Unit Count",
    y = "LOS "
  ) +
  scale_color_discrete(name = "Unit Count") +
  scale_x_continuous(
    breaks = seq(0, max(encounter_unit_switch$UNIT_COUNT), by = 1)) +
  scale_y_continuous(
    breaks = seq(0, max(encounter_unit_switch$LOS_NO_SRC), by = 1)) +
  theme(plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
        axis.title = element_text(size = 12),
        axis.text = element_text(size = 10),
        strip.text = element_text(size = 12, face = "bold"),
        panel.spacing = unit(1.5, "lines"),
        legend.position = "none")
```

### Full Population - LOS Index Distribution
```{r population visualization}
ggplot(encounter_unit_switch, aes(x = UNIT_COUNT, y = LOS_INDEX)) +
  geom_jitter(aes(colour = factor(UNIT_COUNT)),
              width = .2) +
  labs(
    title = "Distribution of Unit Count and LOS Index",
    x = "Unit Count",
    y = "LOS Index"
  ) +
  scale_color_discrete(name = "Unit Count") +
  scale_x_continuous(
    breaks = seq(0, max(encounter_unit_switch$UNIT_COUNT), by = 1)) +
  scale_y_continuous(
    breaks = seq(0, max(encounter_unit_switch$LOS_INDEX), by = 1)) +
  theme(plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
        axis.title = element_text(size = 12),
        axis.text = element_text(size = 10),
        strip.text = element_text(size = 12, face = "bold"),
        panel.spacing = unit(1.5, "lines"),
        legend.position = "none")
```

## Correlations
### LOS Index X Unit Count (no outliers)
```{r los index no outliers, message=FALSE}
# remove outliers
no_outliers <- encounter_unit_switch %>%
  filter(UNIT_COUNT <= 10,
         LOS_INDEX <= 10) %>%
  mutate(UNIT_COUNT_F = factor(UNIT_COUNT, levels = 1:10, ordered = TRUE)) %>%
  arrange(UNIT_COUNT_F)

# calc median performance by unit count grouping
median_data <- no_outliers %>%
  group_by(UNIT_COUNT_F) %>%
  reframe(
    Median_LOS = median(LOS_INDEX, na.rm = TRUE),
    UNIT_COUNT_NUM = as.numeric(UNIT_COUNT_F)
  )

# group percents 
n_counts <- no_outliers %>%
  group_by(UNIT_COUNT_F) %>%
  summarise(n = n(), .groups = "drop") %>%
  mutate(label = percent(n / sum(n), accuracy = 0.01))

# Main Plot:
plot_main <- ggplot(no_outliers, aes(x = UNIT_COUNT_F, y = LOS_INDEX)) + 
  geom_violin(aes(fill = UNIT_COUNT_F), trim = TRUE, alpha = 0.6, color = "gray30", show.legend = FALSE) +
  geom_boxplot(width = 0.1, outlier.shape = NA, fill = "white", color = "black") +
  geom_smooth(data = median_data, aes(x = UNIT_COUNT_NUM, y = Median_LOS), method = "lm", se = FALSE, linewidth = 1.5, color = "#FF00004D") +
  geom_hline(yintercept = 1, linetype = "dashed", color = "black", alpha = 0.3, linewidth = 1) +
  scale_y_continuous(breaks = seq(0, 10, by = 1), limits = c(0, 10)) +
  labs(
    title = "LOS Index by Unit Count",
    subtitle = "Unit Count <= 10; LOS Index <= 10",
    y = "LOS Index",
    x = NULL
  ) +
  theme(plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.y = element_text(size = 14),
    axis.text.x = element_text(size = 14,), 
    axis.ticks.x = element_line(linewidth = 0.8),
    panel.grid.minor = element_blank(),
    plot.margin = margin(t = 10, r = 10, b = -10, l = 10))

# Table Plot: 
plot_table <- ggplot(n_counts, aes(x = UNIT_COUNT_F, y = 1)) +
  geom_text(aes(label = label), size = 4, color = "black", fontface = "plain") + 
  scale_y_continuous(expand = c(0,0)) +
  labs(y = "%") + 
  theme_void() +
  theme(
    axis.title.y = element_text(size = 12, face = "plain", angle = 0, vjust = 0.5),
    plot.margin = margin(t = 0, r = 10, b = 10, l = 10)
  )

# 4. Combine
final_plot <- plot_main / plot_table + plot_layout(heights = c(10, 1))
final_plot
```

### LOS Index X Unit Count (no elective)
```{r los index no surgery, message=FALSE}
# remove outliers
no_outliers <- encounter_unit_switch %>%
  filter(UNIT_COUNT <= 10,
         LOS_INDEX <= 10,
         ADMIT_TYPE_DESC_SRC != "ELECTIVE") %>%
  mutate(UNIT_COUNT_F = factor(UNIT_COUNT, levels = 1:10, ordered = TRUE)) %>%
  arrange(UNIT_COUNT_F)

# calc median performance by unit count grouping
median_data <- no_outliers %>%
  group_by(UNIT_COUNT_F) %>%
  reframe(
    Median_LOS = median(LOS_INDEX, na.rm = TRUE),
    UNIT_COUNT_NUM = as.numeric(UNIT_COUNT_F)
  )

# group percents 
n_counts <- no_outliers %>%
  group_by(UNIT_COUNT_F) %>%
  summarise(n = n(), .groups = "drop") %>%
  mutate(label = percent(n / sum(n), accuracy = 0.01))

# Main Plot:
plot_main <- ggplot(no_outliers, aes(x = UNIT_COUNT_F, y = LOS_INDEX)) + 
  geom_violin(aes(fill = UNIT_COUNT_F), trim = TRUE, alpha = 0.6, color = "gray30", show.legend = FALSE) +
  geom_boxplot(width = 0.1, outlier.shape = NA, fill = "white", color = "black") +
  geom_smooth(data = median_data, aes(x = UNIT_COUNT_NUM, y = Median_LOS), method = "lm", se = FALSE, linewidth = 1.5, color = "#FF00004D") +
  geom_hline(yintercept = 1, linetype = "dashed", color = "black", alpha = 0.3, linewidth = 1) +
  scale_y_continuous(breaks = seq(0, 10, by = 1), limits = c(0, 10)) +
  labs(
    title = "LOS Index by Unit Count",
    subtitle = "Unit Count <= 10; LOS Index <= 10",
    y = "LOS Index",
    x = NULL
  ) +
  theme(plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.y = element_text(size = 14),
    axis.text.x = element_text(size = 14,), 
    axis.ticks.x = element_line(linewidth = 0.8),
    panel.grid.minor = element_blank(),
    plot.margin = margin(t = 10, r = 10, b = -10, l = 10))

# Table Plot: 
plot_table <- ggplot(n_counts, aes(x = UNIT_COUNT_F, y = 1)) +
  geom_text(aes(label = label), size = 4, color = "black", fontface = "plain") + 
  scale_y_continuous(expand = c(0,0)) +
  labs(y = "%") + 
  theme_void() +
  theme(
    axis.title.y = element_text(size = 12, face = "plain", angle = 0, vjust = 0.5),
    plot.margin = margin(t = 0, r = 10, b = 10, l = 10)
  )

# 4. Combine
final_plot <- plot_main / plot_table + plot_layout(heights = c(10, 1))
final_plot
```

### LOS Index X Unit Count (no outliers level of care)
```{r no outliers LOC, message=FALSE}
# remove outliers
no_outliers <- encounter_unit_switch %>%
  filter(FLAG == TRUE,
         UNIT_COUNT <= 10,
         LOS_INDEX <= 10) %>%
  mutate(UNIT_COUNT_F = factor(UNIT_COUNT, levels = 2:10, ordered = TRUE)) %>%
  arrange(UNIT_COUNT_F)

# calc median performance by unit count grouping
median_data <- no_outliers %>%
  group_by(UNIT_COUNT_F) %>%
  reframe(
    Median_LOS = median(LOS_INDEX, na.rm = TRUE),
    UNIT_COUNT_NUM = as.numeric(UNIT_COUNT_F)
  )

# group percents 
n_counts <- no_outliers %>%
  group_by(UNIT_COUNT_F) %>%
  summarise(n = n(), .groups = "drop") %>%
  mutate(label = percent(n / sum(n), accuracy = 0.01))

# Main Plot:
plot_main <- ggplot(no_outliers, aes(x = UNIT_COUNT_F, y = LOS_INDEX)) + 
  geom_violin(aes(fill = UNIT_COUNT_F), trim = TRUE, alpha = 0.6, color = "gray30", show.legend = FALSE) +
  geom_boxplot(width = 0.1, outlier.shape = NA, fill = "white", color = "black") +
  geom_smooth(data = median_data, aes(x = UNIT_COUNT_NUM, y = Median_LOS), method = "lm", se = FALSE, linewidth = 1.5, color = "#FF00004D") +
  geom_hline(yintercept = 1, linetype = "dashed", color = "black", alpha = 0.3, linewidth = 1) +
  scale_y_continuous(breaks = seq(0, 10, by = 1), limits = c(0, 10)) +
  labs(
    title = "LOS Index by Unit Count",
    subtitle = "Unit Count <= 10; LOS Index <= 10",
    y = "LOS Index",
    x = NULL
  ) +
  theme(plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.y = element_text(size = 14),
    axis.text.x = element_text(size = 14,), 
    axis.ticks.x = element_line(linewidth = 0.8),
    panel.grid.minor = element_blank(),
    plot.margin = margin(t = 10, r = 10, b = -10, l = 10))

# Table Plot: 
plot_table <- ggplot(n_counts, aes(x = UNIT_COUNT_F, y = 1)) +
  geom_text(aes(label = label), size = 4, color = "black", fontface = "plain") + 
  scale_y_continuous(expand = c(0,0)) +
  labs(y = "%") + 
  theme_void() +
  theme(
    axis.title.y = element_text(size = 12, face = "plain", angle = 0, vjust = 0.5),
    plot.margin = margin(t = 0, r = 10, b = 10, l = 10)
  )

# 4. Combine
final_plot <- plot_main / plot_table + plot_layout(heights = c(10, 1))
final_plot
```

### LOS Index Level of Care
```{r flag comp, message=FALSE}
# remove outliers
no_outliers <- encounter_unit_switch %>%
  filter(UNIT_COUNT <= 10,
         LOS_INDEX <= 10) %>%
  mutate(UNIT_COUNT_F = factor(UNIT_COUNT, levels = 1:10, ordered = TRUE)) %>%
  arrange(UNIT_COUNT_F)

ggplot(no_outliers, aes(x = LOS_INDEX, fill = FLAG)) +
  geom_density(alpha = 0.6) + 
  geom_vline(xintercept = 1, linetype = "dashed", color = "black", alpha = 0.3, linewidth = 1) +
  facet_grid(FLAG ~ ., 
             labeller = as_labeller(c("FALSE" = "Standard", "TRUE" = "Non-Standard"))) +
  labs(
    title = "Level of Care Transfer Effect on LOS Index",
    x = "LOS Index",
    y = "Density"
  ) +
  theme(
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    strip.text = element_text(size = 14, face = "bold"),
    strip.background = element_rect(fill = "gray95", color = "gray80"),
    panel.grid.minor = element_blank(),
    legend.position = "none"
  )

no_outliers %>%
  group_by(FLAG) %>%
  summarise(count = n(),
            med = median(LOS_INDEX),
            avg = mean(LOS_INDEX),
            sd = sd(LOS_INDEX),
            `los_index_<=1` = mean(LOS_INDEX <= 1, na.rm = TRUE) * 100) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                font_size = 10,
                full_width = F)
```

## Boarders
### Boarder Data
```{r boarder data, message=FALSE}
# load ed boarder data exclude MSSN
ed_boarders <- read_csv("ed_boarder.csv") %>%
  filter(!(`ARRV DEPT (group)` %in% c("MSSN OS", "MSSN LB"))) %>%
  select(CSN, `Admit Decision to Depart(Admits & Boarder)`) %>%
  mutate(BOARD_DURATION = case_when(
    is.na(`Admit Decision to Depart(Admits & Boarder)`) ~ 0,
    TRUE ~ `Admit Decision to Depart(Admits & Boarder)`))

# get strictly emergency encouters
emergency_encounters <- encounter_unit_switch %>%
  filter(ADMIT_TYPE_DESC_SRC == "EMERGENCY")

# join IPCAP with ED boarding data
ip_boarding_join <- emergency_encounters %>%
  left_join(ed_boarders, by = c("EPIC_CSN" = "CSN"))

# quality check for missing encounters
boarding_qc <- ip_boarding_join %>% filter(is.na(BOARD_DURATION))
nrow(boarding_qc)/nrow(ip_boarding_join) * 100

# remove nas
ip_boarding_final <- ip_boarding_join %>%
  filter(!is.na(BOARD_DURATION))

```

### Boarding Distribution
```{r boarding distribution}
no_outliers <- ip_boarding_final %>%
  filter(BOARD_DURATION <= 7500,
         LOS_INDEX <= 10)

ggplot(no_outliers, aes(x = BOARD_DURATION, y = LOS_INDEX)) +
  geom_hex(bins = 50) +
  geom_smooth(method = "gam", color = "red") +
  scale_fill_viridis_c()
```

### LOS Index Boarders
```{r boarder comparison, message=FALSE}
# We manually define the sequence and append Inf (Infinity)
breaks_seq <- c(seq(0, 1500, by = 500), Inf)

# Create labels that are easy to read (e.g., "0-250", ..., "1500+")
labels_seq <- c("0-500 min", "500-1000 min", "1000-1500 min", "1500+ min")

no_outliers_bins <- ip_boarding_final %>%
  filter(BOARD_DURATION <= 7500,
         LOS_INDEX <= 5) %>%
  mutate(board_bin = cut(BOARD_DURATION, 
                         breaks = breaks_seq, 
                         labels = labels_seq,
                         right = FALSE))

ggplot(no_outliers_bins, aes(x = LOS_INDEX)) +
  geom_density(fill = "firebrick", alpha = 0.5, color = "darkred") +
  geom_vline(xintercept = 1, linetype = "dashed", color = "black", linewidth = 0.8) +
  facet_wrap(~ board_bin, ncol = 1) + 
  theme_minimal() +
  labs(
    title = "System LOS Index Density by Boarding Duration",
    subtitle = "Boarding <= 7500 min, LOS_INDEX <= 5",
    x = "LOS Index",
    y = "Density"
  ) +
  theme(
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    strip.text = element_text(size = 14, face = "bold"),
    strip.background = element_rect(fill = "gray95", color = "gray80"),
    panel.grid.minor = element_blank(),
    legend.position = "none"
  )

hospitals <- unique(no_outliers_bins$FACILITY_MSX)
for (hosp in hospitals) {
  p <- ggplot(no_outliers_bins %>%
           filter(FACILITY_MSX == hosp), aes(x = LOS_INDEX)) +
  geom_density(fill = "firebrick", alpha = 0.5, color = "darkred") +
  geom_vline(xintercept = 1, linetype = "dashed", color = "black", linewidth = 0.8) +
  facet_wrap(~ board_bin, ncol = 1) + 
  theme_minimal() +
  labs(
    title = paste(hosp, "LOS Index Density by Boarding Duration"),
    subtitle = "Boarding <= 7500 min, LOS_INDEX <= 5",
    x = "LOS Index",
    y = "Density"
  ) +
  theme(
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    strip.text = element_text(size = 14, face = "bold"),
    strip.background = element_rect(fill = "gray95", color = "gray80"),
    panel.grid.minor = element_blank(),
    legend.position = "none"
  )
  print(p)
}

no_outliers_bins %>%
  group_by(FACILITY_MSX, board_bin) %>%
  summarise(count = n(),
            med = median(LOS_INDEX),
            avg = mean(LOS_INDEX),
            sd = sd(LOS_INDEX),
            `los_index_<=1` = mean(LOS_INDEX <= 1, na.rm = TRUE) * 100) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                font_size = 10,
                full_width = F)
```

# Correlation Tests
### Full Population - LOS Index Correlation
```{r correlation tests, warning=FALSE}
# pearson correlation test
corr_test_pearson <- cor.test(
  x = encounter_unit_switch$UNIT_COUNT,
  y = encounter_unit_switch$LOS_INDEX,
  method = "pearson"
)

# spearman correlation test
corr_test_spearman <- cor.test(
  x = encounter_unit_switch$UNIT_COUNT,
  y = encounter_unit_switch$LOS_INDEX,
  method = "spearman"
)
print(corr_test_pearson)
print(corr_test_spearman)
```

### By DRG - LOS Index Correlation
```{r drg correlation}
# Correlation study by DRG
drg <- encounter_unit_switch %>%
  group_by(MSDRG_CD_SRC) %>%
  summarise(n_encounters = n()) %>%
  filter(n_encounters > 100,
         !is.na(MSDRG_CD_SRC)) %>%
  pull(MSDRG_CD_SRC)

DRG_correlation <- lapply(drg, function(d) {
  df_subset <- encounter_unit_switch %>%
    filter(MSDRG_CD_SRC == d)
  
  # Filter out NA values for the two variables we care about
  df_complete <- df_subset %>%
    filter(!is.na(UNIT_COUNT) & !is.na(LOS_NO_SRC))
  
  n_complete <- nrow(df_complete)
  
  # check for sufficient (3) complete observations
  if (n_complete < 3) { 
    return(data.frame(
      N_Encounters = n_complete,
      R_Value = NA,
      P_Value = NA,
      Error_Reason = paste("Not enough finite observations (N < 3). N =", n_complete)
    ))
  }
  
  # Check for variance within DRG group
  if (sd(df_complete$UNIT_COUNT) == 0 || sd(df_complete$LOS_INDEX) == 0) {
    return(data.frame(
      N_Encounters = n_complete,
      R_Value = NA,
      P_Value = NA,
      Error_Reason = "Zero variance in one or both variables"
    ))
  }
  
  # execute the correlation test using the clean data
  result <- tryCatch(
    {
      corr_test <- cor.test(
        x = df_complete$UNIT_COUNT,
        y = df_complete$LOS_INDEX
      )
      
      data.frame(
        N_Encounters = n_complete,
        R_Value = corr_test$estimate,
        P_Value = corr_test$p.value,
        Error_Reason = "Success"
      )
    },
    error = function(e) {
      data.frame(
        N_Encounters = n_complete,
        R_Value = NA,
        P_Value = NA,
        Error_Reason = paste("Error:", as.character(e))
      )
    }
  )
  return(result)
})

# Combine the results into a single data frame
correlation_summary <- data.frame(
  MSDRG_CD_SRC = drg,
  do.call(rbind, DRG_correlation)
)

final_summary_table <- correlation_summary %>%
  left_join(
    encounter_unit_switch %>% ungroup() %>% distinct(MSDRG_CD_SRC, MSDRG_DESC_MSX),
    by = "MSDRG_CD_SRC") %>%
  # Filter out the groups where the correlation could not be calculated
  filter(Error_Reason == "Success") %>%
  mutate(
    # Format p-value for readability
    P_Value = format.pval(P_Value, digits = 3, eps = 0.001),
    # Determine significance
    Is_Significant = P_Value < 0.05,
    # Classify the strength and direction
    Effect = case_when(
      Is_Significant & R_Value > 0.3 ~ "Strong/Moderate Positive Effect",
      Is_Significant & R_Value < -0.3 ~ "Strong/Moderate Negative Effect",
      Is_Significant ~ "Weak Significant Effect",
      TRUE ~ "Not Significant"
    )
  ) %>%
  # Select and arrange final columns
  select(
    MSDRG_CD_SRC, MSDRG_DESC_MSX, N_Encounters, R_Value, 
    P_Value, Is_Significant, Effect
  ) %>%
  arrange(desc(R_Value))

head(final_summary_table, 20) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                font_size = 10,
                full_width = F)
```